---
title: "Overweight or Obesity in Northeast Ohio's Children"
author: "Better Health Partnership Children's Health Initiative"
date: "Report 4 (released March 2019) describing children ages 2-18 in July 2017 - June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warnings = FALSE)
```

```{r packages_list, include=FALSE}
#read in shape files
library(rgdal)

library(maptools)

#leaflet tool
library(leaflet)

#standard R packages
library(readxl); library(here)
library(janitor); library(tidyverse)
```


```{r load_bhp_data}
# load in bhp data aggregated by zip code
zip_df <- read_xlsx(here("data", "cleaned_counties14_by_zip_2019-02-18.xlsx")) %>%
  clean_names()
```

```{r load_zcta_layers}
# load in zcta layers
zcta_layers <- readOGR(dsn = "data", layer = "cb_2017_us_zcta510_500k", verbose = FALSE)
```

```{r identify_layers_for_mapping}
# select zcta layers that are included in the zip_df data
zcta_use <- subset(zcta_layers, ZCTA5CE10 %in% unique(zip_df$zip))

zcta_use_union <- unionSpatialPolygons(zcta_use, zcta_use$ZCTA5CE10)

zcta_use_map <- fortify(zcta_use_union) 

zcta_use_map_w_data_df <- zcta_use_map %>%
  mutate(id=as.numeric(id)) %>%
  left_join(zip_df,by=c("id"="zip"))
```

```{r set_zoom_options_for_leaflet}
#zoom options
minZoom_val = 9
maxZoom_val = 14

#lat long data needed to specify leaflet boundaries 
min_long = min(zcta_use_map_w_data_df$long)
max_long = max(zcta_use_map_w_data_df$long)
min_lat = min(zcta_use_map_w_data_df$lat)
max_lat = max(zcta_use_map_w_data_df$lat)

long_buffer=0.1*(max_long-min_long)
lat_buffer=0.1*(max_lat-min_lat)

```

```{r create_polyFunc}

polyFunc<-function(groupname, dat){
    poly<-dat %>%
        filter(id==groupname) %>% 
        select(long, lat)
    return(Polygons(list(Polygon(poly)), groupname))
}

```


```{r specify_obesity_numerator_denominator_categories}
zcta_use_data_obesity <- zcta_use_map_w_data_df %>%
  distinct(id, # specify the variables you want to include
           kids_overweight_or_obese, # here
           kids_with_bmi, # here
           prop_overweightorobese) %>%  # and here
  mutate(prop_overweightorobese_c = 
             case_when(prop_overweightorobese<=0.299 ~ 1,
                       prop_overweightorobese<=0.349 ~ 2,
                       prop_overweightorobese<=0.399 ~ 3,
                       prop_overweightorobese> 0.399 ~ 4),
         prop_overweightorobese_c = 
             factor(prop_overweightorobese_c,
                    levels= c("1","2","3","4"),
                    labels = c("Less than 30%",
                               "30 up to 35%",
                               "35 up to 40%",
                               "40% or More")))
```

```{r specify_polygons_for_obesity}
zcta_use_obesity_ids <- zcta_use_data_obesity$id

##base R approach
polygons_obesity<-lapply(zcta_use_obesity_ids,
                 function(x) polyFunc(x, dat=zcta_use_map_w_data_df)) 

sp.polygon_obesity <- SpatialPolygons(polygons_obesity)

df.polygon_obesity <- SpatialPolygonsDataFrame(sp.polygon_obesity, 
                                     data=data.frame(row.names=zcta_use_obesity_ids, zcta_use_data_obesity))
df.polygon_obesity <- df.polygon_obesity[order(df.polygon_obesity$prop_overweightorobese),]
```

```{r generate_map_obesity}
popup <- ifelse(!is.na(df.polygon_obesity$prop_overweightorobese),
          paste0("Zip Code: ", df.polygon_obesity$id, "<br>", 
                                "% of kids overweight or obese: ", round(df.polygon_obesity$prop_overweightorobese*100,1), "<br>",
                "# of kids overweight or obese: ", df.polygon_obesity$kids_overweight_or_obese, "<br>",
                "Total # of kids with BMI: ", df.polygon_obesity$kids_with_bmi),
          paste0("Zip Code: ", df.polygon$id, "<br>", 
                "Less than 100 Children Reported"))

pal <- colorFactor(
    palette = "YlOrRd",
#    palette = "magma",
    domain = df.polygon_obesity$prop_overweightorobese_c[!is.na(df.polygon_obesity$prop_overweightorobese_c)])

map_zcta_obesity <- leaflet() %>% 
  #specify type of map to have in background
  addProviderTiles("OpenStreetMap.Mapnik",
                   options = providerTileOptions(minZoom = minZoom_val, maxZoom = maxZoom_val)
                   ) %>%
  #add first layer of polygons - here I am laying down the filled zip codes 
  addPolygons(data = subset(df.polygon_obesity, !is.na(prop_overweightorobese_c)), 
              fillColor = ~pal(prop_overweightorobese_c), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.5, # can specify how transparent we want the fill to be
              weight = 0.3, # can specify how thick we want the borders of regions to be
              smoothFactor = 0.2
              #popup = popup
              ) %>%
  #add second layer of polygons - here I am laying down the borders of the different tracts
  addPolygons(data = df.polygon_obesity, 
              fillOpacity = 0.0, #want fill to be zero because I am only interested in the borders
              weight = 0.8, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = 
                df.polygon_obesity$prop_overweightorobese_c[!is.na(df.polygon_obesity$prop_overweightorobese_c)], 
            position = "bottomright", 
            title = "% of Children With<br>Obesity:") %>%
  setMaxBounds(lng1=max_long+long_buffer,
               lng2=min_long-long_buffer,
               lat1=min_lat-lat_buffer,
               lat2=max_lat+lat_buffer) %>%
    addMiniMap(toggleDisplay = TRUE)
```

```{r, fig.height = 7.5, fig.width = 9.5}
map_zcta_obesity
```

We report here on the care of 199,571 Ohio children ages 2-18, from July 2017 through June 2018 for whom complete height and weight information are available. 

Overall, **33.3%** of these children had a Body-Mass Index in the 85th percentile or higher, and are thus classified either as overweight or obese. 

- The map is shaded to indicate obesity rates for zip codes reporting at least 100 children. 
- The map describes the 14 Ohio counties which reported on at least 4,000 children.
